---
title: "TidyTuesday: Scooby Doo"
output: html_notebook
---

Source theme function

```{r setup}
source("../R/theme_mark.R")

library(lubridate)
library(here)
```
Make a Scooby Doo palette for the characters (based on [this picture](https://camo.githubusercontent.com/87b411c4ff874f7af1469cc38512cd3dce2844a6a50b65f501dd540209f2d05a/68747470733a2f2f6e79706f73742e636f6d2f77702d636f6e74656e742f75706c6f6164732f73697465732f322f323031392f30392f74762d73636f6f62792d646f6f2d32622e6a70673f7175616c6974793d39302673747269703d616c6c26773d36313826683d3431302663726f703d31))

```{r}
scooby_pal = c("velma" = "#f0a200",
               "shaggy" = "#b1c300",
               "scooby" = "#ae7900",
               "fred" = "#009ad3",
               "daphnie" = "#6b4091")

scooby_pal_ext = c(scooby_pal, "#ffdb6f", "#af7c91", "#a7e3fa", "#f29b78", "#83d18e")
```

Load the data

```{r load-data}

tuesdata <- tidytuesdayR::tt_load('2021-07-13')
```


```{r glimpse-df}
scooby_df = tuesdata$scoobydoo

glimpse(scooby_df)
```

Clean up a few variables that came in as the wrong data type:
- If we're comparing ratings or engagement, we will want `imdb` and `engagement` to be numeric
- If we are looking at any of the booleans (e.g. `non_suspect`, `arrested`) to be boolean rather than character
- Convert `jeepers`, etc. to numerics too

```{r}
scooby_clean = scooby_df %>% 
  mutate(across(c(imdb, engagement), as.numeric),
         across(where(~is_character(.x) && any(str_detect(.x, "TRUE|FALSE"))), 
                as.logical),
         across(jeepers:rooby_rooby_roo, .fns = ~as.numeric(na_if(., "NULL"))))


scooby_clean %>% 
  glimpse()

```


How much Scooby Doo is there? 
Let's see the number of episodes for each season of each show.

```{r}
scooby_clean %>% 
  count(series_name, season)
```

A lot!

Okay first thing I want to see is how the different series ratings differ.
I'm going to remove movies, crossovers, and specials for consistency.

```{r scooby-tv}
scooby_tv = scooby_clean %>% 
  filter(str_detect(season, "Crossover|Movie|Special", negate = T))
```

What does the rating variable look like?

```{r}
scooby_tv %>% 
  count(imdb)
```

And if we convert this to numeric?

```{r}
scooby_clean %>% 
  filter(!is.na(imdb)) %>% 
  group_by(series_name, season) %>% 
  mutate(episode_num = row_number(), .before = "title") %>% 
  select(index, date_aired, series_name, season, episode_num, imdb)
```


Throw that into a plot really quick!

```{r}
scooby_tv %>% 
  ggplot(., aes(x = index, y = imdb))+
  geom_smooth(colour = my_col_pal[3], alpha = .8, se=T, na.rm=T, span = .25, fill = my_col_pal[4])+
  geom_point(aes(colour = series_name), size = 1)+
  scale_colour_discrete(guide = guide_none())+
  scale_y_continuous(name = NULL, limits = c(0,10), expand = expansion())+
  scale_x_continuous(name = "Episode #", expand = expansion(add = c(5, 15)))+
  # scale_x_date(date_breaks = scales::label_date_short())+
  theme_mark(md=T, plots_pane = T)
```


Label some of the series:
- First season
- Best Season
- Lowest season

```{r}
series_ratings_labels = scooby_tv %>% 
  group_by(series_name) %>% 
  summarise(
    year = min(year(date_aired)),
    median_date = median(date_aired),
    first_episode = min(index),
    median_episode = median(index),
    avg_rating = mean(imdb, na.rm=T),
    rating_sd = sd(imdb, na.rm=T),
    episodes = n(),
    .groups = "drop"
  ) %>% 
  mutate(note = case_when(avg_rating == min(avg_rating) ~ "**Lowest avg. rating**<br>",
                          avg_rating == max(avg_rating) ~ "**Highest avg. rating**<br>",
                          first_episode == min(first_episode) ~ "**First season**<br>"),
         label_x = if_else(median_episode == min(median_episode), median_episode+60, median_episode),
         label_y = case_when(avg_rating == min(avg_rating) ~ min(avg_rating)-2,
                          avg_rating == max(avg_rating) ~ max(avg_rating)+1.8,
                          T ~ avg_rating+1.5),
         series_label = str_c(coalesce(note, ""),
                              series_name,
                              "<br>(",year, ")<br>", 
                              "<br>*Mean = ", round(avg_rating, 1), 
                              "; SD = ", round(rating_sd, 1), "*"),
         label_flag = (first_episode == min(first_episode) | 
           avg_rating == min(avg_rating) | avg_rating == max(avg_rating)))

series_ratings_labels
  
```

Add these labels to our plot

```{r}
scooby_ratings = scooby_tv %>% 
    ggplot(., aes(x = index, y = imdb))+
    # Ratings trend line
    geom_smooth(colour = my_col_pal[3], alpha = .8, se=T, na.rm=T, span = .2, fill = my_col_pal[4])+
    # Non-highlighted series
    geom_point(data = filter(scooby_tv, !(series_name %in% series_ratings_labels[series_ratings_labels$label_flag, "series_name", drop=T])), 
               colour = my_col_pal[3], size = 3, alpha = .7)+
    # Highlighted series
    geom_point(data = filter(scooby_tv, series_name %in% series_ratings_labels[series_ratings_labels$label_flag, "series_name", drop=T]), 
               aes(colour = series_name), 
               size = 3, alpha = .9)+
  # Textboxes
    ggtext::geom_textbox(data=filter(series_ratings_labels, label_flag),
                         aes(label = series_label, x = label_x, y = label_y),
                         family = fonts$text, 
                         text.colour = my_col_pal[2], box.colour = my_col_pal[3], size = 4, 
                         minwidth = unit(7, "cm"), maxwidth = unit(9, "cm"), minheight = unit(2, "cm"))+
  # Scales
    scale_colour_manual(values = my_col_pal[5:7], guide = guide_none())+
    scale_y_continuous(name = "IMDB Rating", breaks = c(0,2.5,5,7.5,10), limits = c(0, NA), expand = expansion(add = c(0,1)))+
    scale_x_continuous(name = "Episode #", expand = expansion(add = c(20, 25)))

scooby_ratings+
    # scale_x_date(date_breaks = scales::label_date_short())+
    labs(title = "Zoinks! That's a lot of mysteries gang", 
         caption = source_caption("ScoobyPedia"))+
    theme_mark(md=T, plots_pane = F, base_size = 16)


# ggsave(filename = here("2021-07-13-ScoobyDoo", "ScoobyDooRatings.png"), 
#        # plot = scooby_ratings,
#        device = "png", 
#        width = 16, height = 9, units = "cm", scale = 3.5, dpi = 320)
```




How about engagement?

```{r}
scooby_engagement = scooby_tv %>% 
    ggplot(., aes(x = index, y = engagement))+
    # Ratings trend line
    geom_smooth(colour = my_col_pal[3], alpha = .8, se=T, na.rm=T, span = .2, fill = my_col_pal[4])+
    # Non-highlighted series
    geom_point(data = filter(scooby_tv, !(series_name %in% series_ratings_labels[series_ratings_labels$label_flag, "series_name", drop=T])), 
               colour = my_col_pal[3], size = 3, alpha = .7)+
    # Highlighted series
    geom_point(data = filter(scooby_tv, series_name %in% series_ratings_labels[series_ratings_labels$label_flag, "series_name", drop=T]), 
               aes(colour = series_name), 
               size = 3, alpha = .9)+
    scale_colour_manual(values = my_col_pal[5:7], guide = guide_none())+
    scale_y_continuous(name = "No. of Ratings", limits = c(0, NA), expand = expansion(add = c(0, 50)))+
    scale_x_continuous(name = "Episode #", expand = expansion(add = c(20, 25)))

scooby_engagement +
    labs(title = "But only two series have stood out amongst the crowd",
         subtitle = '"Scooby Doo, Where Are You!" (1969) and "Scooby-Doo Mystery Incorporated" (2010) are the only series with more than 300 reviews for a single episode',
         caption = source_caption("ScoobyPedia"))+
    theme_mark(md=T, plots_pane = F, base_size = 14)

# ggsave(filename = here("2021-07-13-ScoobyDoo", "ScoobyDooRatings-Engagement.png"), 
#        # plot = scooby_ratings,
#        device = "png", 
#        width = 16, height = 9, units = "cm", scale = 3.5, dpi = 320)
```

```{r}
library(patchwork)
# Patch these together

(scooby_ratings + scooby_engagement)* theme_mark(md=T, base_size = 14) + 
  patchwork::plot_annotation(theme = theme_mark(md=T, base_size = 16), caption = source_caption("Scoobypedia"), 
                             title = "Zoinks! That's a lot of mysteries gang",
                             subtitle = "While Scooby Doo's TV series are consistently rated quite well, two series in particular attract a lot more attention than the others.")+
  plot_layout(widths = c(6,4))

# ggsave(filename = here("2021-07-13-ScoobyDoo", "ScoobyDooRatings-Combined.png"), 
#        # plot = scooby_ratings,
#        device = "png", 
#        width = 16, height = 9, units = "cm", scale = 3.5, dpi = 320)
```





Who gets caught/does the capturing/unmasking the most?

```{r}
scooby_tv %>% 
    select(matches("^(caught|captured|unmask)")) %>% 
    pivot_longer(everything(), names_to = c("what", "who"), names_sep = "_", values_to = "caught") %>% 
    count(who, what, wt = caught, name = "times") %>% 
    ggplot(., aes(x = times, y = who, fill = who))+
    geom_col()+
  scale_fill_manual(values = set_names(c(scooby_pal_ext[c(1:5, 7)], "#9c9496"), c(names(scooby_pal), "not", "other")))+
    facet_wrap(~what, nrow=1)+
    theme_mark(plots_pane=T)
```

How do phrases change over time?

```{r}
scooby_tv %>% 
  # group_by(series_name, season) %>%
  group_by(year = year(date_aired)) %>% 
  summarise(date = median(date_aired),
            across(c(jeepers, jinkies, zoinks, groovy), sum, na.rm=T)) %>%
  # select(date_aired, jeepers, jinkies, zoinks, groovy) %>%
  pivot_longer(cols = jeepers:groovy, names_to = "phrase", values_to = "times") %>% 
  ggplot(., aes(x = date, y = times, group = phrase))+
  # ggplot(., aes(x = date_aired, y = times, group = phrase))+
  geom_smooth(aes(colour = phrase), size = 1.25, fill = my_col_pal[4], span = .2)+
  # geom_jitter(aes(colour = phrase), alpha = .3)+
  scale_colour_manual(values = set_names(scooby_pal[c(1,2,4,5)], c("jinkies", "zoinks", "groovy", "jeepers")))+
  scale_x_date(name = NULL, breaks = scales::breaks_pretty(n=8), labels = scales::label_date_short(format = c("%Y", "%b", "%d")), expand = expansion())+
  scale_y_continuous(name = NULL, breaks = scales::breaks_pretty(), limits = c(0, NA), expand = expansion())+
  theme_mark(plots_pane = T)
```








# Discarded

Plot each episode's use of phrases

```{r}
scooby_tv %>% 
  # group_by(series_name, season) %>%
  # summarise(date = median(date_aired),
  #           across(c(jeepers, jinkies, zoinks, groovy), sum, na.rm=T)) %>%
  select(date_aired, jeepers, jinkies, zoinks, groovy) %>%
  pivot_longer(cols = jeepers:groovy, names_to = "phrase", values_to = "times") %>% 
  # ggplot(., aes(x = date, y = times, group = phrase))+
  ggplot(., aes(x = date_aired, y = times, group = phrase))+
  # geom_smooth(aes(colour = phrase), size = 1.25, fill = my_col_pal[4])+
  geom_jitter(aes(colour = phrase), alpha = .7)+
  scale_colour_manual(values = set_names(scooby_pal[c(1,2,4,5)], c("jinkies", "zoinks", "groovy", "jeepers")))+
  scale_x_date(name = NULL, breaks = scales::breaks_pretty(n=8), labels = scales::label_date_short(format = c("%Y", "%b", "%d")), expand = expansion())+
  scale_y_continuous(name = NULL, breaks = scales::breaks_pretty(), limits = c(0, NA), expand = expansion())+
  theme_mark(plots_pane = T)
```

So how did Scooby Doo change from 1969 to 2010?

```{r}
scooby_changes = scooby_tv %>% 
  filter(series_name %in% c("Scooby Doo, Where Are You!", "Scooby-Doo Mystery Incorporated")) %>% 
  separate_rows(monster_name, monster_gender, monster_type, monster_subtype,monster_species, sep = ",") %>% 
  group_by(series_name) %>% 
  mutate(year = min(year(date_aired)),
         series_label = factor(str_c(series_name, "\n(", year, ")"))) %>% 
  ungroup() %>% 
  count(series_name, monster_type)


  

scooby_tv %>% 
  # Keep only the two most popular
  filter(series_name %in% c("Scooby Doo, Where Are You!", "Scooby-Doo Mystery Incorporated")) %>% 
  # Separate multiple monsters
  separate_rows(monster_name, monster_gender, monster_type, monster_subtype,monster_species, sep = ",") %>% 
  group_by(series_name) %>% 
  mutate(year = min(year(date_aired)),
         series_label = factor(str_c(series_name, "\n(", year, ")"))) %>% 
  ungroup() %>% 
  # Count the number of monster types from each series
  count(series_name, monster_type, year) %>%
  # Pivot to make a row for each monster type
  pivot_wider(id_cols = c(monster_type), names_from = "year", names_prefix = "series_", values_from = "n") %>%
  replace_na(list(series_2010 = 0, series_1969 = 0)) %>%
  # Make a midpoint and difference to plot the arrows
  mutate(midpoint = (series_2010+series_1969)/2,
         diff = series_2010-series_1969,
         monster_type = factor(monster_type)) %>%
  arrange(-diff) %>% 
  mutate(monster_type = fct_inorder(monster_type)) %>% 
  ggplot(., aes(y = reorder(monster_type, diff)))+
  # Found the idea here to get the arrow in the middle:
  # https://community.rstudio.com/t/how-to-draw-an-arrow-in-the-middle-of-the-segment/7305
  geom_segment(aes(x = midpoint, xend = series_2010, yend = monster_type))+
  geom_segment(aes(x = series_1969, xend = midpoint, yend = monster_type),
               arrow = arrow(type = "closed", length = unit(.25, "cm"), ends = "last"), arrow.fill=my_col_pal[3])+
  geom_point(aes(x = series_1969), colour = scooby_pal[1])+
  geom_point(aes(x = series_2010), colour = scooby_pal[4])+
  scale_y_discrete()+
  # Gave up because I couldn't immediately get a legend to work and concluded this may not even be that interesting
  # scale_colour_manual(values = set_names(c("Scooby Doo, Where Are You!", "Scooby-Doo Mystery Incorporated"),
  #                                        scooby_pal[c(1,4)]))
  theme_mark(md=T)+
  theme(legend.position = "bottom")
  
```



