---
title: "R Notebook"
output: html_notebook
---

# Get the Data

```{r}
library(extrafont)
library(sf)
library(raster)
library(tidyverse)
library(spData)
# note that spDataLarge needs to be installed via:
# install.packages("spDataLarge", repos = "https://nowosad.github.io/drat/", type = "source")
library(spDataLarge) 

# Visualization packages
library(tmap)
library(here)
source(here("R", "theme_mark.R"))
```

# Geographic data in R

**Vector data model** uses points, lines, and polygons.
High precision, not necessarily high accuracy.
Typically used in social sciences for its discrete borders.

**Raster data model** divides the surface into cells of equal size. 
Web mapping use raster data for background images.
Aggregate spatial features to given *resolution* => they are consistent over space and can scale.
Typically used in environmental science because of their abundance of remote sensing data.

## Vector data

Based on points defined on a coordinate reference system (CRS).
These points can be longitude/latitude on a geographic CRS, or in the case of London, Easting/Northing which comes from the British National Grid (a system of geographic references used in Great Britain).

# Making maps with R

[Link to chapter](https://geocompr.robinlovelace.net/adv-map.html)

```{r}
map_nz = tm_shape(nz)+
  tm_polygons()

map_nz +
  tm_shape(nz_elev) + 
  tm_raster(alpha = 0.7)
```

Making NZ water

```{r}
nz_water = st_union(nz) %>% st_buffer(22200) %>% 
  st_cast(to = "LINESTRING")

nz_water
```

```{r}
map_nz+
  # Plot elevation
  tm_shape(nz_elev)+
  tm_raster(alpha = .7)+
  # Plot water
  tm_shape(nz_water)+
  tm_lines()+
  # Plot high points
  tm_shape(nz_height)+
  tm_dots()
```

`tmap` accepts aesthetic arguments directly

```{r}
tm_shape(nz)+
  tm_fill(col = viz_colours[1], alpha = .6)+
  tm_borders(col = my_col_pal[2])
```
`tmap` can use continuous data for colours via column reference in the form of a string.

```{r}
tm_shape(nz)+
  tm_fill(col = "Population", alpha = .6, 
          palette = viz_colours[c(3,1,6,5)])+
  tm_borders(col = my_col_pal[2])
```

We can use `breaks` or `n` to set the size of bins

```{r}
tm_shape(nz)+
  tm_fill(col = "Population", alpha = .6, 
          palette = viz_colours[c(3,1,6,5,8)],
          breaks = c(0, 100000, 250000, 500000, 1000000, 2000000))+
  tm_borders(col = my_col_pal[2])
```
Or we can use algorithms to specify the break styles with the `style` argument

```{r}
map(c("pretty", "equal", "quantile", "jenks", "cont", "cat"),
    function(selected_style){
      tm_shape(nz)+
        tm_fill(col = "Population", alpha = .6, 
                style = selected_style,
                title = selected_style)+
        tm_borders(col = my_col_pal[2])
      })
```

While we have been using `tm_fill()+tm_borders()`, it looks like `tm_polygons()` achieves both.

```{r}
tm_shape(nz) + 
  tm_polygons("Population", palette = "YlOrBr")
```

We can add layout elements with `tm_layout()`

```{r}
map_nz+
  tm_layout(title = "Title",
            frame = F)
```

# Read in MID and MIDLOC Data

## Country Codes

List of country codes used in Correlates of War data.

```{r}
cc_file = here("data", "MID5", "COW country codes.csv")

cc = read_csv(cc_file)

glimpse(cc)
```

## Militarized Interstate Dispute (MID)

>Version 5 of the Militarized Interstate Dispute (MID) data collection compiled by the Correlates of War Project provides information about conflicts in which one or more states threaten, display, or use force against one or more other states between 1816 and 2014. By definition, “Militarized interstate disputes are united historical cases of conflict in which the threat, display or use of military force short of war by one member state is explicitly directed towards the government, official representatives, official forces, property, or territory of another state. Disputes are composed of incidents that range in intensity from threats to use force to actual combat short of war” (Jones et al. 1996: 163).

MIDA data, which contains data on MIDs from 1816-2014, at the dispute level, i.e. one record per militarized dispute.

```{r}
mida_file = here("data", "MID5", "MIDA 5.0.csv")

mida = read_csv(mida_file)

glimpse(mida)
```

MIDB data, which contains data on MIDs from 1816-2014, at the dispute level, i.e. one record per militarized dispute participant.

```{r}
midb_file = here("data", "MID5", "MIDB 5.0.csv")

midb = read_csv(midb_file) %>% 
  left_join(cc, by = c("stabb" = "StateAbb", "ccode" = "CCode"))

glimpse(midb)
```
```{r}
midb_aggressors = filter(midb, sidea == 1)

glimpse(midb_aggressors)
```




## Militarized Interstate Dispute Location (MIDLOC)

>The Militarized Interstate Dispute Location (MIDLOC) dataset details the precise geographic location of Militarized Interstate Disputes (MIDs) in the post-Napoleonic era (from 1816 - 2010), complementing the Correlates of War (COW) Project's considerable empirical data collection.

```{r}
midloca_file = here("data", "MIDLOC_2.1", "MIDLOCA_2.1.csv")

midloca = read_csv(midloca_file)

glimpse(midloca)
```

There seem to be a fair few disputes where the locations are unknown

```{r}
midloca %>% 
  filter(if_any(c(midloc2_xlongitude,midloc2_ylatitude), is.na))
```
Because we're concerned with mapping these disputes, let's subset the data to disputes with known locations. 

```{r}
midloca_known = midloca %>% 
  filter(!if_any(c(midloc2_xlongitude,midloc2_ylatitude), is.na))
```

Now let's have a look at the distribution of disputes over time.

```{r}
ggplot(midloca_known)+
  geom_histogram(aes(x = year), colour = my_col_pal[1])+
  scale_y_continuous(name = "MIDs", expand = expansion(add = .5))+
  theme_mark(plots_pane=T)
```


Group MIDs by time period: 1800-1850, 1850-1900, 1900-1950, 1950-2000, 2000-2010.

```{r}
midloca_known_grouped = midloca_known %>% 
  mutate(time_period = case_when(between(year, 1800, 1850) ~ "1800-1850",
                                 between(year, 1851, 1900) ~ "1851-1900",
                                 between(year, 1901, 1950) ~ "1901-1950",
                                 between(year, 1951, 2000) ~ "1951-2000",
                                 T ~ "2001-2010"))

midloca_known_grouped %>% 
  count(time_period)
```


```{r}
library(sp)
world_sp = as(world, Class = "Spatial") # from an sf object to sp
# sp functions ...
world_sf = st_as_sf(world_sp)           # from sp to sf
```


```{r}
midloca_sf = st_as_sf(midloca_known_grouped, 
             coords = c("midloc2_xlongitude", "midloc2_ylatitude"),
             crs = 4326)

glimpse(midloca_sf)
```

```{r}
tm_shape(world_sf)+
  tm_borders(col = my_col_pal[2])+
  tm_shape(midloca_sf)+
  tm_dots()+
  # tm_facets(along = "year", free.coords = F)+
  tm_layout(frame = F, 
            bg.color = my_col_pal[1], 
            fontfamily = fonts$body, 
            title.fontfamily = fonts$title, 
            main.title = "Militarized Interstate Disputes (1816-2010)")
```

```{r}
library(gganimate)

ggplot(midloca_sf)+
  geom_sf(data = world, 
          colour = my_col_pal[2], fill = my_col_pal[1])+
  geom_sf(colour = viz_colours[2])+
  labs(title = "Military Interstate Disputes {year}")+
  theme_mark(plots_pane = T)+
  theme(panel.grid.major = element_blank())+
  transition_manual(year)
  
```

