---
title: "R Notebook"
output: html_notebook
---

# Get the Data

```{r}
library(sf)
library(raster)
library(tidyverse)
library(spData)
# note that spDataLarge needs to be installed via:
# install.packages("spDataLarge", 
# repos = "https://nowosad.github.io/drat/", type = "source")
library(spDataLarge) 

# Visualization packages
library(tmap)
library(here)
source(here("R", "theme_mark.R"))
```

# Geographic data in R

**Vector data model** uses points, lines, and polygons.
High precision, not necessarily high accuracy.
Typically used in social sciences for its discrete borders.

**Raster data model** divides the surface into cells of equal size. 
Web mapping use raster data for background images.
Aggregate spatial features to given *resolution* => they are consistent over space and can scale.
Typically used in environmental science because of their abundance of remote sensing data.

## Vector data

Based on points defined on a coordinate reference system (CRS).
These points can be longitude/latitude on a geographic CRS, or in the case of London, Easting/Northing which comes from the British National Grid (a system of geographic references used in Great Britain).

# Making maps with R

[Link to chapter](https://geocompr.robinlovelace.net/adv-map.html)

```{r}
map_nz = tm_shape(nz)+
  tm_polygons()

map_nz +
  tm_shape(nz_elev) + 
  tm_raster(alpha = 0.7)
```

Making NZ water

```{r}
nz_water = st_union(nz) %>% st_buffer(22200) %>% 
  st_cast(to = "LINESTRING")

nz_water
```

```{r}
map_nz+
  # Plot elevation
  tm_shape(nz_elev)+
  tm_raster(alpha = .7)+
  # Plot water
  tm_shape(nz_water)+
  tm_lines()+
  # Plot high points
  tm_shape(nz_height)+
  tm_dots()
```

`tmap` accepts aesthetic arguments directly

```{r}
tm_shape(nz)+
  tm_fill(col = viz_colours[1], alpha = .6)+
  tm_borders(col = my_col_pal[2])
```
`tmap` can use continuous data for colours via column reference in the form of a string.

```{r}
tm_shape(nz)+
  tm_fill(col = "Population", alpha = .6, 
          palette = viz_colours[c(3,1,6,5)])+
  tm_borders(col = my_col_pal[2])
```

We can use `breaks` or `n` to set the size of bins

```{r}
tm_shape(nz)+
  tm_fill(col = "Population", alpha = .6, 
          palette = viz_colours[c(3,1,6,5,8)],
          breaks = c(0, 100000, 250000, 500000, 1000000, 2000000))+
  tm_borders(col = my_col_pal[2])
```
Or we can use algorithms to specify the break styles with the `style` argument

```{r}
map(c("pretty", "equal", "quantile", "jenks", "cont", "cat"),
    function(selected_style){
      tm_shape(nz)+
        tm_fill(col = "Population", alpha = .6, 
                style = selected_style,
                title = selected_style)+
        tm_borders(col = my_col_pal[2])
      })
```



